Concurrent Programming with CXL

### 2020-05-16 book/paper.tex
Inserting a chapter
More book stuff

==== Safety violation ====
__main__ 0 [0,12-29]     30 { 'count': 0, 'done': { 1: False, 2: False } }
up       1 [1-5]          6 { 'count': 0, 'done': { 1: False, 2: False } }
up       2 [1-10]        11 { 'count': 1, 'done': { 1: False, 2: True  } }
up       1 [6-10]        11 { 'count': 1, 'done': { 1: True,  2: True  } }
__main__ 0 [30-37,39-44] 44 { 'count': 1, 'done': { 1: True,  2: True  } }
\end{verbatim}
\end{code}
\caption{The output of running Figure~\ref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

@@

A CXL machine has the following components:
\begin{itemize}
\item Code:  This is an immutable and finite list of CXL machine instructions,
generated from a CXL program.  The types of instructions will be described later.
\item Shared memory: A CXL machine has just one memory location containing
a CXL.
\item Processes:  Initially there is just the main process, but any process
can spawn an unbounded number of other processes and processes may terminate.
Each process has an immutable tag, a program counter, a stack of CXL values,
and a single mutable general purpose \emph{register} that contains a CXL.
\end{itemize}

The state of a process is called a \emph{context}: it contains the value of
its program counter, stack, and register.  The state of the CXL machine
consists of the value of its memory and the multiset (or \emph{bag}) of
contexts.  It is a multiset of contexts because two different processes can
be in the same state.

@@

\chapter{Naive Attempts at a Solution}

@@

\begin{figure}
\begin{code}
\begin{verbatim}
==== Livelock ====
__main__ 0 [0,35-63]    64 {'flags': {2: False, 1: False}, 'count': 0,
                                'done': {2: False, 1: False}}
up       1 [1-13,21-24] 25 {'flags': {2: False, 1: True}, 'count': 0,
                                'done': {2: False, 1: False}}
\end{verbatim}
\end{code}
\caption{Error output of Figure~\ref{fig:upflags}.}
\label{fig:upflagsoutput}
\end{figure}

>>>>

==== Safety violation ====
__main__ () [0,12-29]     30 { 'count': 0, 'done': { 1: False, 2: False } }
up       1  [1-5]          6 { 'count': 0, 'done': { 1: False, 2: False } }
up       2  [1-10]        11 { 'count': 1, 'done': { 1: False, 2: True  } }
up       1  [6-10]        11 { 'count': 1, 'done': { 1: True,  2: True  } }
__main__ () [30-37,39-44] 44 { 'count': 1, 'done': { 1: True,  2: True  } }
\end{verbatim}
\end{code}
\caption{The output of running Figure~\ref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

@@

A CXL machine has the following components:
\begin{itemize}
\item Code:  This is an immutable and finite list of CXL machine instructions,
generated from a CXL program.  The types of instructions will be described later.
\item Shared memory: A CXL machine has just one memory location containing
a CXL.
\item Processes:  Initially there is just the main process, but any process
can spawn an unbounded number of other processes and processes may terminate.
Each process has an immutable \emph{name tag} (specified when spawning the
process), a program counter, a stack of CXL values, and a single mutable
general purpose \emph{register} that contains a CXL.
\end{itemize}

The state of a process is called a \emph{context}: it contains the value of
its name tag, program counter, stack, and register.
The state of the CXL machine
consists of the value of its memory and the multiset (or \emph{bag}) of
contexts.  It is a multiset of contexts because two different processes can
be in the same state.
The initial state of the CXL memory is \texttt{dict\{\}}.
The context bag has an initial context in it with name tag
\texttt{dict\{ .name: "\_\_main\_\_", .tag: () \}}, pc 0,
register \texttt{dict\{\}}, and an empty stack.
Each machine instruction updates the state in some way.
Figure~\ref{fig:incstate} shows an example of a reachable state for 
the program in Figure~\ref{fig:inc}.

@@

\chapter{Critical Sections}

\begin{figure}
\begin{code}
\verbatiminput{cs.cxl}
\end{code}
\caption{A barebones critical section.}
\label{fig:cs}
\end{figure}

A critical section is often modeled as processes in an infinite loop
entering and exiting the critical section.  In CXL:

\begin{code}
\begin{verbatim}
while True:
    # enter critical section
    @cs: pass;  # in critical section
    # exit critical section
;
\end{verbatim}
\end{code}

Here \texttt{@cs} is a \emph{label}, identifying a location in the
CXL machine code.  The first thing we need to ensure is that there
cannot be two processes at the critical section.  We would like to
place an assertion at the \texttt{@cs} label that specifies that
only the current process can be there.  CXL in fact supports this.
It has a method \texttt{atLabel($L$)}, where $L$ is the string name
of a label (in this case, \texttt{"cs"}).  The method returns a
bag of name tags of processes executing at that label.  The bag is
represented by a dictionary that maps each element in the bag to the
number of times the element appears in the bag.

Figure~\ref{fig:cs} shows the code.
The assertion also makes use of the method \texttt{nametag()} that
returns the name tag of the current process.
If you run the code through CXL, the assertion should fail because
there is no code yet for entering and exiting the critical section.

\chapter{Naive Attempts at Implementing Critical Sections}

@@

\begin{figure}
\begin{code}
\begin{verbatim}
==== Livelock ====
__main__ () [0,35-63]    64 {'flags': {2: False, 1: False}, 'count': 0,
                                'done': {2: False, 1: False}}
up       1  [1-13,21-24] 25 {'flags': {2: False, 1: True}, 'count': 0,
                                'done': {2: False, 1: False}}
\end{verbatim}
\end{code}
\caption{Error output of Figure~\ref{fig:upflags}.}
\label{fig:upflagsoutput}
\end{figure}
